---
title: "Distribution of tuna fisheries near Tristan da Cunha"
author: "Cian Luck"
date: "24 Nov 2021"
output:
  html_notebook:
  df_print: default
highlight: pygments
toc: yes
toc_float:
  toc_collapsed: true
toc_depth: 2

editor_options:
  chunk_output_type: inline
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Background

The marine protected zone (MPZ) of Tristan da Cunha is the Atlantic's largest no-take area, providing an important conservation area for several marine species, including the endangered Tristan albatross, the southern bluefin tuna, and the near threatened blue shark. While longline fishing is prohibited within the MPZ, longline fishing vessels are extremely active in adjacent waters and often fish right up to the boundary of Tristan da Cunha's waters. The activity of these fisheries has important implications for management of many of the species that Tristan da Cunha aims to conserve. Understanding the scale and distribution of longline fishing effort is vital to understanding the pressures that this activity might be placing on key species of interest, but equally important is idnetifying how this activity is likely to change over time.

In this analysis we will:

1. provide a detalied overview of fishing activity by longline vessels in the waters adjacent to Tristan da Cunha
2. explore multi-year trends in longliner activity
3. compare longline fishing activity to publicly available data sourced from regional fisheries management organisations.
  
# Setup

```{r setup, include = FALSE}
# set working directory to parent folder
knitr::opts_knit$set(root.dir = normalizePath("..")) 
```

```{r, include=FALSE}
# run script with custom ggplot theme functions
source("analysis/theme_gfw modified funs - cian.R", local = knitr::knit_global())
```

Load packages: 

```{r, message=FALSE}
library(tidyverse)
library(bigrquery)
library(DBI)
library(fishwatchr)
library(glue)
library(lubridate)
library(here)
library(sf)
library(extrafont)
library(patchwork)
library(terra)
```

Establish connection to Big Query project:

```{r}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), 
                      project = "world-fishing-827", 
                      use_legacy_sql = FALSE)
```

Load detailed shapefile of Tristan da Cunha:

```{r}
tdc_sf <- st_read("geodata/tdc_osm/tristan_da_cunha_archipelago_osm.shp")
st_crs(tdc_sf) <- 4326
```

# Overview of longline fishing activity

This query selects all fishing activity by longline vessels near Tristan da Cunha, and includes in which fishing area (if any) the fishing took place:

```{r}
query_1 <- readr::read_file(str_c("queries", "q_tdc_tuna_fishing_effort_aoi.sql", sep="/"))
```

Run the query using `fishwatchr::gfw_query()`

```{r}
ll_aoi <- fishwatchr::gfw_query(query = query_1,
                            run_query = TRUE,
                            con = con)$data
```

Save the query output to local device. Load from here for future analysis, without needing to rerun query above.

```{r}
saveRDS(ll_aoi, "data_production/data/longliner_dist_2016_2021_aoi.rds")
ll_aoi <- read_rds("data_production/data/longliner_dist_2016_2021_aoi.rds")
```

How many longline vessels belonging to each flag were active in the north and south fishing areas each year between 2016 and 2021.

First, create a dataframe that sums the fishing hours (`fishing_hours`) and number of longline vessels (`n_vessels`) per year, flag, and fishing area:

```{r}
ll_vessels_flag_by_year <- ll_aoi %>% 
  mutate(year = date %>% lubridate::year()) %>% 
  group_by(year, best_flag, aoi) %>% 
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE),
            n_vessels = n_distinct(ssvid),
            .groups = "keep")
```

Create a barplot of the number of longline active vessels per year, in the **north** fishing area, coloured by vessel flag:

```{r}
# create a summary table of the top 5 most active flags
# useful for plotting
top_flag_n <- ll_aoi %>% 
  filter(aoi == "north") %>% 
  group_by(best_flag) %>% 
  summarise(n_vessels = n_distinct(ssvid)) %>% 
  mutate(prop = n_vessels/sum(n_vessels)) %>% 
  top_n(5, wt = n_vessels) %>%
  arrange(desc(n_vessels)) 

# create a barplot of number of vessels per year, coloured by vessel flag
# save as an object called p_fv_n
p_fv_n <- ll_vessels_flag_by_year %>% 
  filter(aoi == "north") %>% 
  # if the flag is not one of the top 5 most active flags, label it as "other"
  mutate(flag_alt = if_else(best_flag %in% top_flag_n$best_flag, best_flag, "other") %>% 
           factor(levels = c(top_flag_n$best_flag, "other"))) %>% 
  ggplot() +
  geom_col(aes(x = year, y = n_vessels, fill = flag_alt)) +
  scale_fill_manual(values = c(gfw_palette("chart")[1:5], "grey50"),
                    name = "Flag",
                    labels = c("TPE", "JPN", "ESP", "KOR", "CHN", "other")) +
  scale_x_continuous(breaks = c(2016, 2017, 2018, 2019, 2020, 2021)) +
  labs(subtitle = "North",
       x = "Year",
       y = "Active longliners") + 
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_blank(),
        plot.subtitle = element_text(size = 14),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 16),
        legend.position = "right")
```

Create a barplot of the number of longline active vessels per year, in the **south** fishing area, coloured by vessel flag:

```{r}
# create a summary table of the top 2 most active flags
# useful for plotting
top_flag_s <- ll_aoi %>% 
  filter(aoi == "south") %>% 
  group_by(best_flag) %>% 
  summarise(n_vessels = n_distinct(ssvid)) %>% 
  mutate(prop = n_vessels/sum(n_vessels)) %>% 
  top_n(2, wt = n_vessels) %>%
  arrange(desc(n_vessels))

# create a barplot of number of vessels per year, coloured by vessel flag
# save as an object called p_fv_s
p_fv_s <- ll_vessels_flag_by_year %>% 
  filter(aoi == "south") %>% 
  # if the flag is not one of the top 5 most active flags, label it as "other"
  mutate(flag_alt = if_else(best_flag %in% top_flag_s$best_flag, best_flag, "other") %>% 
           factor(levels = c(top_flag_s$best_flag, "other"))) %>% 
  ggplot() +
  geom_col(aes(x = year, y = n_vessels, fill = flag_alt)) +
  scale_fill_manual(values = c(gfw_palette("chart")[c(1,3)],"grey50"),
                    name = "Flag") +
  scale_x_continuous(breaks = c(2016, 2017, 2018, 2019, 2020, 2021)) +
  labs(subtitle = "South",
       x = "Year",
       y = "Active longliners") + 
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        plot.subtitle = element_text(size = 14),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 16),
        legend.position = "right")
```

Combine the two plots into one using `patchwork`

```{r, warning=FALSE, fig.width=7, fig.height=5}
(p_fv_n / p_fv_s) + 
  plot_annotation(title = "Number of longline vessels active per year", 
                  theme = theme_gfw_cian() + theme(plot.title = element_text(size = 16)))
```

# Trends in longliner activity over time

Here we want to look at how fishing activity by longline vessels has changed in the north and south fishing areas, between 2016 and 2021. We also want to include sea surface temperature (SST) data to look at how these trends might correlate with SST. We already have a monthly summary of fishing activity in `ll_monthly_summary`, so next we need to load SST data, and in this analysis we are using data sourced from the [Aqua Modis](https://neo.gsfc.nasa.gov/view.php?datasetId=MYD28M) project.

Here, we use the `terra` package to read in raster files of monthly average SST data near Tristan da Cunha

```{r}
# read in raster filenames
file_names_sst <- list.files(path = "data_production/data/rasters/aqua_modis_sst_float_tiff", pattern = "*.tiff", full.names = TRUE)

# load these files as a single multi-layer raster
sst_rast <- terra::rast(file_names_sst) 

# change the names of each raster layer to the year and month (format: YYYY-MM)
names(sst_rast) <-  names(sst_rast) %>% str_sub(8, -7)

# Replace outlier temperatures (>100°C) with NA
sst_rast[sst_rast >= 100] <- NA
```

Save raster stack for future use (See [this](https://geocompr.robinlovelace.net/read-write.html#raster-data-write) from Lovelace et al for most apropriate datatype):

```{r}
sst_rast %>% writeRaster(filename = "data_production/data/rasters/sst_2016_2021_multiyear.tiff",
                         datatype = "INT4S", overwrite = FALSE)
sst_rast <- terra::rast("data_production/data/rasters/sst_2016_2021_multiyear.tiff")
```

Create spatial objects of the fishing areas (also referred to as areas of interest, coded as `aoi_` in this case) using the simple features `sf` package. This will allow us to extract relevant SST data for each area:

```{r}
# create a geometry object including the dimensions of the northern fishing area
aoi_geom_n <- list(rbind(c(-20,-30), c(10,-30), c(10,-40), c(-20,-40), c(-20,-30))) %>% 
  st_polygon() %>% 
  st_sfc(crs = 4326)

# convert this to a simple features dataframe with a single data column (id = "north")
aoi_sf_n <- data.frame(id = "north") %>% 
  st_sf(geometry = aoi_geom_n)

# create a geometry object including the dimensions of the southern fishing area
aoi_geom_s <- list(rbind(c(-20,-40), c(15,-40), c(15,-50), c(-20,-50), c(-20,-40))) %>% 
  st_polygon() %>% 
  st_sfc(crs = 4326)

# convert this to a simple features dataframe with a single data column (id = "south")
aoi_sf_s <- data.frame(id = "south") %>% 
  st_sf(geometry = aoi_geom_s)

# bind the two sf dataframes together
aoi_sf <- rbind(aoi_sf_n, aoi_sf_n)
```

Extract summary of sst for each fishing area including:

* minimum
* mean
* standard deviation (sd)
* maximum

```{r}
# north fishing area
sst_north_summary <- sst_rast %>% 
  terra::extract(vect(aoi_sf_n)) %>% # extract all sst values within north fishing area 
  group_by(ID) %>% 
  summarise(across(names(sst_rast), list(min = ~min(.x, na.rm = TRUE),  # summarise sst data
                                              mean = ~mean(.x, na.rm = TRUE), 
                                              sd = ~sd(.x, na.rm = TRUE),
                                              max = ~max(.x, na.rm = TRUE)))) %>% 
  pivot_longer(cols = -ID) %>% # the rest of the code reformats the data into our preferred structure
  separate(name, into = c("date", "name"), sep = "_") %>% 
  pivot_wider(names_from = name,
              values_from = value) %>% 
  mutate(aoi = "north") %>% 
  dplyr::select(date, aoi, min, mean, sd, max)

# south fishing area
sst_south_summary <- terra::extract(sst_rast, vect(aoi_sf_s)) %>% 
  group_by(ID) %>% 
  summarise(across(names(sst_rast), list(min = ~min(.x, na.rm = TRUE), 
                                              mean = ~mean(.x, na.rm = TRUE), 
                                              sd = ~sd(.x, na.rm = TRUE),
                                              max = ~max(.x, na.rm = TRUE)))) %>% 
  pivot_longer(cols = -ID) %>%
  separate(name, into = c("date", "name"), sep = "_") %>% 
  pivot_wider(names_from = name,
              values_from = value) %>% 
  mutate(aoi = "south") %>% 
  dplyr::select(date, aoi, min, mean, sd, max)

# rbind both summary dataframes together
sst_summary <- rbind(sst_north_summary, sst_south_summary)
```

Save output locally and load from here for future use:

```{r}
sst_summary %>% saveRDS("data_production/data/sst_summary_longliner_aoi_2016_2021.rds")
sst_summary <- read_rds("data_production/data/sst_summary_longliner_aoi_2016_2021.rds")
```

Format data to code `date` as a datetime object, then extract `month` and `year`

```{r}
sst_summary <- sst_summary %>% 
  mutate(date = date %>% lubridate::ym(),
         month = date %>% lubridate::month(),
         year = date %>% lubridate::year())
```

Merge the monthly summaries of longline fishing activity with monthly summaries of SST:

```{r}
ll_sst_month <- merge(x = ll_monthly_summary, 
                      y = sst_summary %>% 
                        # rename columns to avoid confusion in combined dataframe
                        dplyr::select(month = date, aoi, sst_min = min, sst_mean = mean, sst_sd = sd, sst_max = max), 
                      # merge by month and area of interest (aoi)
                      by = c("month", "aoi"), all.x = TRUE)
```

Create a combined plot of monthly fishing hours, number of active vessels, and mean (+/- standard deviation) SST in the north and south fishing areas. We'll do this by creating three separate plots (vessel hours, number of vessels, SST), and then stacking these plots together using `patchwork`

```{r, warning=FALSE, fig.width=10, fig.height=7}
# create a set of labels for annotating the plots
aoi_labs <- c("north" = "North", "south" = "South")

# fishing hours
p_hours <- ll_sst_month %>% 
  ggplot() +
  geom_col(aes(x = month, y = fishing_hours), fill = gfw_palette("chart")[1]) +
  scale_x_date(date_breaks = "1 year",
             date_labels = "%Y",
             limits = c(ymd("2016-01-01"), ymd("2021-10-31"))) +
  labs(x = "Date (monthly bins)",
       y = "Fishing hours") + 
  scale_y_continuous(limits = c(0, 20000)) +
  facet_grid(.~aoi,
             labeller = labeller(aoi = aoi_labs)) +
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines"))

# number of vessels
p_fv <- ll_sst_month %>% 
  ggplot() +
  geom_col(aes(x = month, y = n_vessels), fill = gfw_palette("chart")[1]) +
  scale_x_date(date_breaks = "1 year",
             date_labels = "%Y",
             limits = c(ymd("2016-01-01"), ymd("2021-10-31"))) +
  labs(x = "Date (monthly bins)",
       y = "Vessels") + 
  scale_y_continuous(limits = c(0, 45)) +
  facet_grid(.~aoi,
             labeller = labeller(aoi = aoi_labs)) +
  theme_gfw_cian() +
   theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_blank(),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines"))

# SST
p_sst <- ll_sst_month %>% 
ggplot() +
  geom_ribbon(aes(x = month, ymin = sst_mean - sst_sd, ymax = sst_mean + sst_sd),
              fill = gfw_palette("chart")[5], alpha = 0.2) +
  geom_line(aes(x = month, y = sst_mean), colour = gfw_palette("chart")[5]) +
  scale_x_date(date_breaks = "1 year",
             date_labels = "%Y",
             limits = c(ymd("2016-01-01"), ymd("2021-10-31"))) +
  labs(x = "Date",
       y = "SST (°C)"
       ) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_grid(.~aoi,
             labeller = labeller(aoi = aoi_labs)) +
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        strip.text = element_blank(),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines"))

# combine plots
p_hours / p_fv / p_sst + 
  plot_annotation(title = "Fishing activity by longline vessels and sea surface temperature (SST)",
                  theme = theme_gfw_cian() + theme(plot.title = element_text(size = 20)))
```

# Comparisons with RFMO datasets

## CCSBT: southern bluefin tuna catch

At the time of writing, the Commission for the Conservation of Southern Bluefin Tuna (CCSBT) have published annual reported catch data for southern bluefin tuna. We compared the observed trends in declining fishing activity by vessels targeting southern bluefin tuna in the southern fishing area with reporte southern bluefin tuna catch over the same period, in the same area.

Our data show a continued decline in fishing effort by the Southern Bluefin tuna fleet to the southeast of Tristan. It would be interesting to compare the decline in this activity with catch data, to see if catch has declined at the same rate as effort.

Catch data was sourced from the CCSBT data repository, accessed (here)[https://www.ccsbt.org/en/content/sbt-data. Data was downloaded as a csv and uploaded into R.

```{r}
sbt_catch <- read.csv("data_production/data/sbt_catch_ccsbt_1965_2020.csv")
```

Filter these data to only include data from 2016 to 2020 and only within the south fishing area, and tidy up the formatting of date and catach (`tonnes`) data:

```{r}
sbt_catch <- sbt_catch %>% 
  filter(year %>% between(2016, 2020),
         lon %>% between(-20, 15),
         lat %>% between(-50, -40)) %>% 
  mutate(date = str_c(year, month, sep = "-") %>% lubridate::ym(),
         tonnes = tonnes %>% as.numeric())
```

Plot total catch by month:

```{r, warning=FALSE, fig.width=10, fig.height=5}
p_s_catch <- sbt_catch %>% 
  group_by(date) %>% 
  summarise(tonnes = sum(tonnes)) %>% 
  ggplot() +
  geom_col(aes(x = date, y = tonnes), fill = gfw_palette("chart")[1]) +
  scale_x_date(date_breaks = "1 year",
             date_labels = "%Y",
             limits = c(ymd("2016-01-01"), ymd("2020-12-31"))) +
  labs(x = "Date (monthly bins)",
       y = "SBT catch (tonnes)") + 
  # scale_y_continuous(limits = c(0, 20000)) +
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        # axis.text.x = element_blank(),
        # axis.title.x = element_blank(),
        plot.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom")


p_s_hours_2 <- ll_sst_month %>% 
  filter(aoi == "south") %>% 
  ggplot() +
  geom_col(aes(x = month, y = fishing_hours), fill = gfw_palette("chart")[1]) +
  scale_x_date(date_breaks = "1 year",
             date_labels = "%Y",
             limits = c(ymd("2016-01-01"), ymd("2020-12-31"))) + # change the time period to up to 2020
  labs(x = "Date (monthly bins)",
       y = "Fishing hours") + 
  scale_y_continuous(limits = c(0, 20000)) +
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom")

p_s_hours_2 / p_s_catch +
  plot_annotation(title = "Fishing activity by longline vessels and catch of Southern Bluefin Tuna (SBT)",
                  subtitle = "Southern AOI",
                  theme = theme_gfw_cian() + theme(plot.title = element_text(size = 16), plot.subtitle = element_text(size = 14)))
```
## ICCAT: blue shark catch

Tristan da Cunha has been recognised as an important conservation area for blue sharks, a near threatened species which is often caught as incidental bycatch but may also be the target species of certain fisheries. Looking at the catch data reported to ICCAT, we can see that Spanish-flagged longline vessels have caught more blue sharks throughout the Southern Atlantic than any other flag.

Load the data on reported blue shark catch per flag in the South Atlantic, sourced from ICCAT statistical [databases](https://www.iccat.int/en/accesingdb.HTML):

```{r}
bsh_catch <- read.csv("data_production/data/iccat_bsh_catch_atlantic_south_2010_2020.csv")

# reformat data into a usable structure
bsh_catch <- bsh_catch %>% 
  pivot_longer(cols = X2010:X2020,
               names_to = "year",
               values_to = "catch") %>% 
  mutate(year = year %>% str_sub(start = 2) %>% as.numeric())
```

Reduce to only flags that caught the most blue shark:

```{r, warning = FALSE, fig.width=7, fig.height=5}
# create a summary table of the top 10 flags that caught the most blue shark
# useful for plotting
bsh_top_flags <- bsh_catch %>%
  filter(!is.na(flag),
         year %>% between(2016, 2020)) %>% 
  group_by(flag) %>% 
  summarise(catch = sum(catch, na.rm=TRUE)) %>% 
  arrange(desc(catch)) %>% 
  top_n(10)

# create a barplot of total blue shark catch per year per flag
bsh_catch %>% 
  mutate(flag = flag %>% factor(levels = bsh_top_flags$flag)) %>% 
  filter(!is.na(flag),
         year %>% between(2016, 2020)) %>% 
  ggplot() +
    geom_col(aes(x = flag, y = catch), fill = gfw_palette("chart")[1]) +
    scale_x_discrete(labels = c("ESP", "PRT", "BRA", "JPN", "NAM", "TPE", "ZAF", "CIV", "KOR", "BLZ")) +
    labs(x = "Flag",
         y = "BSH catch (tonnes)",
         title = "Blue shark (BSH) catch in the South Atlantic\nreported to ICCAT",
         subtitle = "2016 - 2020") + 
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom")
```

Plot total blue shark catch by Spanish-flagged longline vessels each year (2016-2020) in the South Atlantic:

```{r, fig.width=7, fig.height=5}
bsh_catch %>% 
  filter(flag == "EU-España",
         year %>% between(2016, 2020)) %>% 
  ggplot() +
    geom_col(aes(x = year, y = catch), fill = gfw_palette("chart")[1]) +
    labs(x = "Year",
         y = "BSH catch (tonnes)",
         title = "Blue shark (BSH) catch by Spanish-flagged longline vessels\nin the South Atlantic reported to ICCAT") + 
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom")
```

Plot fishing effort each year (2016-2020) by Spanish-flagged longline vessels in the north fishing area:

```{r}
ll_aoi %>% 
  filter(aoi == "north",
         best_flag == "ESP") %>% 
  mutate(year = date %>% lubridate::year()) %>% 
  group_by(year) %>% 
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE),
            n_vessels = n_distinct(ssvid, na.rm = TRUE),
            .groups = "keep") %>% 
  ggplot() +
  geom_col(aes(x = year, y = fishing_hours), fill = gfw_palette("chart")[1]) +
  labs(x = "Year",
       y = "Fishing hours",
       title = "Fishing effort by Spanish-flagged longline vessels",
       subtitle = "Northern fishing area") + 
  scale_x_continuous(n.breaks = 6) +
  theme_gfw_cian() +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        legend.position = "bottom")
    
```
